Let There Be Light: \\ Early Observations of the Normal Type Ia Supernova \abc\ \\Show No Evidence for a Dark Period

### 2017-06-14 doc/ms.tex
Include major rewrite of the paper to reflect new figures and updated analysis of the explanation for the early observations of iPTF 16abc.

%% If you want to create your own macros, you can do so
%% using \newcommand. Your macros should appear before
%% the \begin{document} command.
%%
\newcommand{\vdag}{(v)^\dagger}
\newcommand\aastex{AAS\TeX}
\newcommand\latex{La\TeX}
\newcommand{\sm}{M_\odot}
\newcommand{\sr}{R_\odot}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%
%% The following section outlines numerous optional output that
%% can be displayed in the front matter or as running meta-data.
%%
%% If you wish, you may supply running head information, although
%% this information may be modified by the editorial offices.
\shorttitle{iPTF16abc}
\shortauthors{Cao et al.}
%%
%% You can add a light gray and diagonal water-mark to the first page 
%% with this command:
\watermark{DRAFT}
%% where "text", e.g. DRAFT, is the text to appear.  If the text is 
%% long you can control the water-mark size with:
%  \setwatermarkfontsize{dimension}
%% where dimension is any recognized LaTeX dimension, e.g. pt, in, etc.
%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

\title{Optifcal Observations of an Extraordinarily Young Type Ia Supernova iPTF16abc}

@@

\author[0000-0002-8036-8491]{Yi Cao}
\affil{eScience Institute and Astronomy Department, University of Washington,
  Seattle, WA 98195}

\author{Friends}
\affil{the intermediate Palomar Transient Factory}

@@

  We report the discovery of an extraordinarily young Type Ia supernova,
  iPTF16abc, by the intermediate Palomar Transient Factory. Our first
  observation was made only $0.18$~d after the onset of the
  supernova explosion. Our analysis shows that the early emission from iPTF16abc is dominated by radioactive decay, and that contributions from supernova shock breakout or the ejecta colliding with a stellar companion are minimal. The combination of (i) a fast initial rise, (ii) the
  early exponential decline in the photospheric velocity, and (iii) the significant presence of ionized carbon in the earliest spectra together provide
  evidence for the strong mixing of radioactive elements in the supernova ejecta. \amiller{There should be something about dark times/lack of dark time for this SN in the abstract.}

@@

%% Keywords should appear after the \end{abstract} command. 
%% See the online documentation for the full list of available subject
%% keywords and the rules for their use.
\keywords{methods: observational --- supernovae: individual (iPTF16abc)}

@@

Although Type Ia supernovae (SNe Ia) have been extensively used as
standardizable candles, their progenitors and explosion
physics are still debated (see a recent review by
\citealt{2014ARA&A..52..107M}). Extremely detailed
observations in the hours to days after explosion are one of the most promising avenues to further
constrain this problem.

While the shock breakout of a SN Ia occurs on a sub-second timescale,
the subsequent quasi-adiabatic expanding and cooling of the unbound
ejecta produces thermal emission that can be used to infer the
original size of the exploding star
\citep{2010ApJ...708..598P,2011ApJ...728...63R}. Comparing models of
this cooling emission to the earliest-phase data of SN~2011fe,
\citet{2012ApJ...744L..17B} concluded that the radius of the SN~2011fe progenitor is $\lesssim0.01\sr$, where $\sr$ is the solar
radius. Combining this size constraint and the measured ejecta mass to
derive the mean density of the progenitor star, we confirmed that the
progenitor star is compact and degenerate\ycao{Ref?}. \ycao{a transition is missing here - not sure how the second half of this paragraph relates to the first} Admittedly, due to the
initial small surface area of the progenitor star, the shock cooling
emission of a SN Ia decays drastically as the ejecta expands. Given
typical parameters for a SN Ia, this thermal emission is visible from
events up to $\sim 10\,\textrm{Mpc}$ within one day of their explosions.

SN Ia resulting from a white dwarf and non-degenerate companion, referred to as the single-degenerate progenitor, early-phase observations may detect excess emission due to the collision of the SN ejecta and the non-degenerate companion \citep{1973ApJ...186.1007W,2010ApJ...708.1025K}. This excess emission was first detected in iPTF14atg,  \citep{2015Natur.521..328C}, a low-velocity SN Ia with a significant and declining ultraviolet pulse detected within a
few days of the SN explosion. This signature is best interpreted as a
SN ejecta-companion collision. Many studies have searched for this excess emission, with most resulting in non-detections
\citep{2010ApJ...722.1691H,2011ApJ...741...20B,2012ApJ...744...38F,
  2012ApJ...744L..17B,2015Natur.521..332O,
  2013ApJ...778L..15Z,2015ApJ...799..106G,2016ApJ...826..144S,
  2015ApJS..221...22I}. An exception is SN2012cg, 
  as a blue excess in its early-phase light curve is attributed to an ejecta-companion collision in \citet{2016ApJ...820...92M} (though see \citealt{2016arXiv161007601S} for an alternative interpretation). 
With only $\sim 10\%$ of single-degenerate progenitors occupying the preferred binary geometry for us to see the collision
signatures, it is not surprising that there are few confirmed detections in the literature \ycao{ref?}.

The detection of SN shock breakout or an ejecta-companion collision is exceedingly rare, and the vast majority of SNe Ia are observed to be 
powered purely by the radioactive decay of $^{56}$Ni. 
SNe Ia may experience a dark period after
the SN shock breakout but before radioactive energy diffuses into
the photosphere depending on how the newly synthesized $^{56}$Ni is mixed and deposited into different layers of the ejecta \citep{2014ApJ...784...85P}. If there is strong
mixing, photons from $^{56}$Ni decay can reach the photosphere rapidly. This leads to a short dark period and a fast initial rise in the observed optical light curve. For weak mixing, it may take several days for
the radioactive energy to diffuse to the photosphere. The initial rise
of the light curve is then moderate \citep{2016ApJ...826...96P}. Thus, the initial rise of SN Ia light curves conveys
information on the distribution of synthesized $^{56}$Ni.

In this paper, we report observations of an extraordinarily young SN
Ia, iPTF16abc, which was discovered by the intermediate Palomar
Transient Factory on 2016 April $3.36$\footnote{All times in this
  paper are in UTC.} at $\textrm{R.A.}=13^h34^m45.49^s$,
$\textrm{Dec.}=+13^d51^m14.3^s$ (J2000) with a $g$-band magnitude of
$21.31\pm0.27$ \citep{2016PASP..128k4502C,2016ATel.8907....1M}. The
transient is spatially coincident with a tidal tail of the galaxy
NGC\,5221, which lies at a distance of $\sim$100\,Mpc. iPTF16abc is not detected to a limit of $g=22.1$\,mag on April $2.42$, less than 1 d prior to discovery. Our spectroscopic follow-up
compaign classified iPTF16abc as a normal SN Ia
\citep{2016ATel.8909....1C}. Our observations and analysis provide multiple lines of evidence that the $^{56}$Ni produced by iPTF16abc was strongly mixed throughout the ejecta. 
\ycao{My personal preference is to not include section by section outline, esp for short papers - this info is clear from section titles. But if you want to include this feel free}

% This paper is organized as follows: Section \ref{sec:obs} describes
% photometric and spectroscopic observations of iPTF16abc. Section
% \ref{sec:usual_staff} establishes that iPTF16abc is a normal SN Ia
% in NGC\,5221. Section \ref{sec:first_light} analyzes the early
% light curve and spectra.

@@

\begin{figure*}[htb]
  \centering
  \includegraphics[width=0.95\textwidth]{lightcurve.pdf}
  \caption{Multi-band light curves of iPTF16abc. Each filter is
  represented with a different color, and each instrument with a 
  different symbol. $t_{max}$ is the time of B-band maximum
    determined by SALT2 (Section \ref{sec:classification}). The solid
    curves are best-fit results from SALT2. The black ticks near the
    top of the figure show epochs of spectroscopic observations.}
  \label{fig:lightcurve}
\end{figure*}

During the spring of 2016, the iPTF survey observed the field of iPTF16abc every night in either the $g$- or $R$-band.
Survey observations were conducted with the
CFH12K camera \citep{2000SPIE.3965...58S} on the Palomar Observatory 48-inch telescope 
 (P48). Images were processed by the IPAC image
subtraction and discovery pipeline which subtracts off the background
galaxy light with stacked pre-SN images and performs forced
point-spread-function (PSF) photometry at the location of the SN.\ycao{Add a reference to Frank's PTFIDE paper - http://iopscience.iop.org/article/10.1088/1538-3873/129/971/014002/pdf} The
photometry is then calibrated to the PTF photometric catalog
\citep{2012PASP..124..854O}.

After discovery, photometric observations in the $g'$, $r'$ and $i'$
filters were obtained with the SED Machine 
(SEDM; \ycao{REF}) mounted on the Palomar Observatory 
60-inch telescope (P60). We utilized the Fremling Automated Pipeline \citep{2016A&A...593A..68F} to subtract galaxy light from the SEDM images using achival Sloan Digital Sky Survey (SDSS) images as reference images. This pipeline then performed forced-PSF photometry at the location of iPTF16abc, which is calibrated to the SDSS catalog \ycao{which DR?}.

photometric observations in the \textit{BVgri} were conducted by the Las Cambres Observatory Global Network (LCOGT) 1-m
telescope network.  PSF photometry was measured on these images using
the \texttt{lcogtsnpipe} pipeline \citep{2016MNRAS.459.3939V}. The
\textit{BV} magnitudes are calibrated to the Fourth USNO CCD
Astrograph Catalog \citep{2013AJ....145...44Z}, and the \textit{gri}
magnitudes are calibrated to SDSS Data Release 6
\citep{2008ApJS..175..297A}.

In space, \textit{Swift} observed iPTF16abc on 14 epochs, beginning $\sim$15 d pre-maximum light through $\sim$5 d post maximum. \ycao{check and confirm numbers} The SN flux is measured via aperture photometry on Ultraviolet-Optical
Telescope (UVOT) images via the usual procedures in the HEASoft, including corrections for the coincident loss and aperture loss. The image counts
are converted to physical fluxes using the latest calibration
\citep{2011AIPC.1358..373B}. There are no pre-SN UVOT images at the SN location available in the \textit{Swift} archive.  Visual inspection of the
UVOT images suggests negligible host-galaxy contamination in our UVOT flux measurements. There is no X-ray emission detected at the 
SN position by the \textit{Swift} X-ray Telescope (XRT).

The multi-color light curves of iPTF16abc are illustrated in 
Figure~\ref{fig:lightcurve}.  For convenience, magnitudes in \textit{all}
filters are in the AB system with a zero point of $3631$\,Jy.

Spectroscopic observations of iPTF16abc were undertaken with a variety
of telescopes and instruments in multiple epochs spanning from a
couple of days after explosion to two months after $B$-band maximum. An
observing log is listed in Table \ref{tab:spec_obs_log}. The spectra were reduced using standard routines in \texttt{IDL}/\texttt{Python}. The optical
spectral evolution of iPTF16abc is illustrated in Figure
\ref{fig:spec_seq}, which excludes high-resolution Very Large Telescope (VLT) spectra for clarity.

\begin{deluxetable*}{cccccc}
  \tablecaption{Spectroscopic observations of iPTF16abc \label{tab:spec_obs_log}}
  \tablehead{
    \colhead{Observation MJD} & \colhead{SN phase} & \colhead{Telescope} &
    \colhead{Instrument} & \colhead{Wavelength Coverage (\AA)}
  }
  \startdata
  $57483.26$ & $-16.4$ & DCT & DeVeny\tablenotemark{1} & $3301$--$7499$ \\
  $57483.88$ & $-15.8$ & Gemini-North & GMOS\tablenotemark{2} & $3800$--$9200$ \\
  $57484.51$ & $-15.1$ & Keck-II & DEIMOS\tablenotemark{3} & $5500$--$8099$ \\
  $57486.51$ & $-13.1$ & Keck-II & DEIMOS\tablenotemark{3} & $5500$--$8099$ \\
  $57488.38$ & $-11.3$ & Keck-I & LRIS\tablenotemark{4} & $3055$--$10411$ \\
  $57489.51$ & $-10.1$ & LCOGT-2m & FLOYDS\tablenotemark{5} & $3301$--$8999$ \\
  $57490.40$ & $ -9.3$ & LCOGT-2m & FLOYDS\tablenotemark{5} & $3301$--$9999$ \\
  $57491.55$ & $ -8.1$ & LCOGT-2m & FLOYDS\tablenotemark{5} & $3300$--$9998$ \\
  $57492.20$ & $ -7.5$ & VLT & X-shooter\tablenotemark{6} & $3300$--$24550$ \\
  $57494.00$ & $ -5.7$ & VLT & UVES\tablenotemark{7} & \\
  $57503.32$ & $ +3.7$ & LCOGT-2m & FLOYDS\tablenotemark{5} & $3300$--$9999$ \\
  $57506.00$ & $ +6.3$ & NOT & ALFOSC\tablenotemark{8} & $3602$--$8098$ \\
  $57508.27$ & $ +8.6$ & LCOGT-2m & FLOYDS\tablenotemark{5} & $3301$--$9999$ \\
  $57518.42$ & $+18.8$ & Keck-I & LRIS\tablenotemark{4} & $3071$--$10208$ \\
  $57520.03$ & $+20.4$ & VLT & X-shooter\tablenotemark{6} & $3300$--$24789$ \\
  $57529.40$ & $+29.8$ & LCOGT-2m & FLOYDS\tablenotemark{5} & $4000$--$8998$ \\
  $57542.41$ & $+42.8$ & LCOGT-2m & FLOYDS\tablenotemark{5} & $4000$--$8998$ \\
  $57550.40$ & $+50.8$ & LCOGT-2m & FLOYDS\tablenotemark{5} & $4001$--$8999$ \\
  $57562.38$ & $+62.7$ & LCOGT-2m & FLOYDS\tablenotemark{5} & $4800$--$9300$ \\
  \enddata
  \tablenotetext{1}{The Deveny Spectrograph \citep{2014SPIE.9147E..2NB}}
  \tablenotetext{2}{The Gemini Multi-Object Sectrograph \citep{2004PASP..116..425H}}
  \tablenotetext{3}{DEep Imaging Multi-Object Spectrograph \citep{2003SPIE.4841.1657F}}
  \tablenotetext{4}{Low-Resolution Imaging Spectrometer \citep{1995PASP..107..375O}}
  \tablenotetext{5}{FLOYDS \url{https://lco.global/observatory/instruments/floyds}}
  \tablenotetext{6}{X-shooter \citep{XShooter}}
  \tablenotetext{7}{Ultraviolet and Visual Echelle Spectrograph \citep{2000SPIE.4008..534D}}
  \tablenotetext{8}{The Andalucia Faint Object Spectrograph and Camera \url{http://www.not.iac.es/instruments/alfosc}}
\end{deluxetable*}

\begin{figure*}[!htb]
  \centering
  \includegraphics[width=1.0\textwidth]{spectra.pdf}
  \caption{Low-resolution spectra of iPTF16abc are shown in the
    chronical orders. In order for better illustration, each spectrum
    is normalized by the median flux value between $6,000$ and
    $7,000\,\textrm{\AA}$ and offset properly.  The phases in units of
    days are noted next to corresponding spectra. Telluric absorption
    bands are grayed out. The narrow \ion{Na}{1}\,D absorption is also
    highlighted in orange.}
  \label{fig:spec_seq}
\end{figure*}

@@

The foreground Galactic extinction toward iPTF16abc
is $E(B-V)=0.0279\,\mathrm{mag}$ \citep{2011ApJ...737..103S}.

Our highest-resolution spectrum of iPTF16abc, obtained with VLT/UVES, shows multiple absorption components for each of the individual lines in the \ion{Ca}{2}\,H$+$K and \ion{Na}{1}\,D doublets (Figure \ref{fig:narrow_features}). This indicates there are two sources of absorption along the line of sight. Fitting
two Gaussian kernels to each line of the \ion{Na}{1} doublet
simultaneously leads to redshifts of $0.02313820\pm0.00000032$ and
$0.02322408\pm0.00000033$. The total \ion{Na}{1}\,D equivalent width (EW) is $0.595\pm0.009\,\textrm{\AA}$ and
$0.609\pm0.008\,\textrm{\AA}$, respectively. Using the empirical
relation between the EW of \ion{Na}{1}\,D lines and
reddening $E(B-V)$ \citep{2012MNRAS.426.1465P}, we derive
$E(B-V)=0.361\pm0.025\textrm{mag}$. We note that this relation has large scatter.

In the VLT/X-shooter spectra of iPTF16abc, we also identify narrow \ion{K}{1}\,7665\,\AA\ and 7699\,\AA\ absorption at the same
redshifts. The double-absorption profiles of \ion{Ca}{2}
and \ion{Na}{1} are not resolved in the VLT/X-shooter spectra. Furthermore, the X-shooter spectra do not show the diffusive
interstellar band at 5780\,\AA. For iPTF16abc, \ion{Na}{1}\,D provides the best estimate of host-galaxy reddening (though see \citealt{2013ApJ...779...38P} for a discussion on why \ion{Na}{1}\,D is a poor tracer of redenning).

>>>>

%% If you want to create your own macros, you can do so
%% using \newcommand. Your macros should appear before
%% the \begin{document} command.
%%
\newcommand{\vdag}{(v)^\dagger}
\newcommand\aastex{AAS\TeX}
\newcommand\latex{La\TeX}
\newcommand{\sm}{M_\odot}
\newcommand{\sr}{R_\odot}
\newcommand{\abc}{iPTF\,16abc}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%
%% The following section outlines numerous optional output that
%% can be displayed in the front matter or as running meta-data.
%%
%% If you wish, you may supply running head information, although
%% this information may be modified by the editorial offices.
\shorttitle{\abc}
\shortauthors{Alphabet Friends}
%%
%% You can add a light gray and diagonal water-mark to the first page 
%% with this command:
\watermark{DRAFT}
%% where "text", e.g. DRAFT, is the text to appear.  If the text is 
%% long you can control the water-mark size with:
%  \setwatermarkfontsize{dimension}
%% where dimension is any recognized LaTeX dimension, e.g. pt, in, etc.
%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

\title{Let There Be Light: \\ Early Observations of the Normal Type Ia Supernova \abc\ \\Show No Evidence for a Dark Period}

@@

% \author[0000-0002-8036-8491]{Yi Cao}
% \affil{eScience Institute and Astronomy Department, University of Washington,
%   Seattle, WA 98195}

\author{The Alphabet Friends}
\affil{the intermediate Palomar Transient Factory}

@@

Early observations of type Ia supernovae (SNe) provide a unique probe of their progenitor systems and explosion physics. Here, we report the intermediate Palomar Transient Factory (iPTF) discovery of an extraordinarily young Type Ia supernova, \abc. The initial detection of the SN was made only $0.18$~d after the time of first light. In the $\sim$24 hr after discovery \abc\ rose by $\sim$2 mag in the $g_\mathrm{PTF}$ band, which is very rare for young type Ia SNe. Furthermore, by measuring the evolution of the photospheric velocity we determine that the time of explosion for \abc\ is approximately equal to the time of first light. Spectra obtained in the week after explosion reveal the presence of strong \ion{C}{2} absorption, which steadily fades over that time period. We show that the unusually fast, near-linear rise in flux for \abc\ cannot be explained by either SN shock breakout and the assciated subsequent cooling or the collision of the SN ejecta with a stellar companion. Instead, we show that each of the unusual characteristics of \abc: (i) the rapid, near-linear rise, (ii) the lack of a dark period between explosion and first light, and (iii) the strong lines from ionized carbon in the earliest spectra, can all be explained if radioactive $^{56}$Ni is strongly mixed throughout the SN ejecta. In the next few years, dozens of very young \textit{normal} SNe Ia will be discovered, and observations similar to those presented here will constrain the white dwarf explosion mechanism.

@@

%% Keywords should appear after the \end{abstract} command. 
%% See the online documentation for the full list of available subject
%% keywords and the rules for their use.
\keywords{methods: observational --- supernovae: individual (\abc)}

@@

Although Type Ia supernovae (SNe Ia) have been extensively used as
standardizable candles, their progenitor systems and explosion
physics are still debated (see a recent review by
\citealt{2014ARA&A..52..107M}). Extremely detailed
observations in the hours to days after explosion are one of the most promising avenues to further
constrain this problem.

While the shock breakout of a SN Ia occurs on a sub-second timescale,
the subsequent quasi-adiabatic expansion and cooling of the unbound
ejecta produces thermal emission that can be used to infer the
radius of the exploding star
\citep{2010ApJ...708..598P,2011ApJ...728...63R}. Comparing models of
this cooling emission to the earliest-phase data of SN~2011fe,
\citet{2012ApJ...744L..17B} concluded that the explosion came from a 
stars with $R_\ast \lesssim 0.02\sr$, where $\sr$ is the solar
radius. Combining the radius constraint with the measured ejecta 
mass \citeauthor{2012ApJ...744L..17B} derive the mean density of the 
progenitor star, confirming that at least some type Ia SNe come from 
compact and degenerate stars.

Early phase observations of SNe Ia from a white dwarf (WD)$+$non-degenarate binary
may detect excess emission, relative to most type Ia SNe, due to the 
collision of the SN ejecta with the non-degenerate companion 
\citep{1973ApJ...186.1007W,2010ApJ...708.1025K}. This excess 
emission was first detected in iPTF\,14atg  \citep{2015Natur.521..328C}, a low-velocity SN Ia with a significant and declining ultraviolet (UV) pulse detected within a
few days of the SN explosion. This UV pulse is best interpreted as a
SN ejecta-companion collision. While such emission requires a 
favorable geometric alignment and is only expected in $\lesssim$ 
10\% of SNe Ia \citep{2010ApJ...708.1025K}, many studies have 
searched for signatures of an ejecta-companion interaction, 
typically resulting in non-detections 
(e.g., \citealt{2010ApJ...722.1691H,2011ApJ...741...20B,2012ApJ...744...38F,
  2012ApJ...744L..17B,2015Natur.521..332O,
  2013ApJ...778L..15Z,2015ApJ...799..106G,2016ApJ...826..144S,
  2015ApJS..221...22I}). A possible exception is SN\,2012cg, 
which exhibited excess blue emission in its early-phase light curve 
\citet{2016ApJ...820...92M} (though see 
\citealt{2016arXiv161007601S} for an interpretation that does not 
invoke ejecta-companion interaction). 
% With only $\sim 10\%$ of single-degenerate progenitors occupying the preferred binary geometry for us to see the collision
% signatures, it is not surprising that there are few confirmed detections in the literature \ycao{ref?}.

The vast majority of SNe Ia are observed 
to be powered purely by the radioactive decay of $^{56}$Ni. 
While the detection of SN shock cooling or ejecta-companion interaction is rare, the level of $^{56}$Ni mixing in the SN ejecta can fundamentally alter the appearance of the SN shortly after explosion. SNe Ia experience a dark period after
the SN shock breakout but before radioactive energy  
diffuses into the photosphere \citep{2014ApJ...784...85P}. The 
duration of this dark period is set by how the newly synthesized 
$^{56}$Ni is mixed and deposited into different layers of the ejecta.
In the case of strong mixing, the dark period is short, or 
non-existent, as photons from $^{56}$Ni decay can reach the 
photosphere rapidly. This also leads to blue optical colors and a 
rapid initial rise in the observed light curve. If the mixing is 
weak and the $^{56}$Ni is confined to the innermost layers of the 
ejecta, the dark period can last for several days as the radioactive 
energy diffuses to the photosphere. The early evolution of such SNe 
results in redder colors and a more moderate rise in luminosity 
\citep{2016ApJ...826...96P}. Thus, the early light curves of even non-exotic type Ia SNe convey information about their progenitor systems by constraining the distribution of synthesized $^{56}$Ni, which in turn constrains the explosion mechanism.

In this paper, we report observations of an extraordinarily young SN
Ia, \abc, which was discovered by the intermediate Palomar
Transient Factory (iPTF) on 2016 April $3.36$\footnote{All times in this
  paper are in UTC.} at $\textrm{R.A.}=13^h34^m45.49^s$,
$\textrm{Dec.}=+13^d51^m14.3^s$ (J2000) with a $g_\mathrm{PTF}$-band magnitude of
$21.31\pm0.27$ \citep{2016ATel.8907....1M}. The
transient is spatially coincident with a tidal tail of the galaxy
NGC\,5221, which lies at a distance of $\sim$100\,Mpc. \abc\ is not detected to a limit of $g=22.1$\,mag on April $2.42$, less than 1 d prior to discovery, and rose by $\sim$2 mag in the 24 hr following its initial detection. Our spectroscopic follow-up
compaign classified \abc\ as a normal SN Ia
\citep{2016ATel.8909....1C}. Our observations and analysis provide multiple lines of evidence that the $^{56}$Ni produced by \abc\ was strongly mixed throughout the ejecta.

% This paper is organized as follows: Section \ref{sec:obs} describes
% photometric and spectroscopic observations of \abc. Section
% \ref{sec:usual_staff} establishes that \abc\ is a normal SN Ia
% in NGC\,5221. Section \ref{sec:first_light} analyzes the early
% light curve and spectra.

@@

\begin{figure*}[htb]
  \centering
  \includegraphics[width=0.95\textwidth]{logLC_with_colors.pdf}
  \caption{Photometric evolution of \abc. \textit{Top}: The $g$-band
  light curve of \abc. Observations from different 
  telescopes are shown with different symbols. Time is measured in 
  rest-frame d relative to the time of first light, $t_0$ (see 
  \S\ref{sec:lc_fit}). Note that the horizonal axis is shown with a linear scale 
  for $-2 \; \mathrm{d} \le t - t_0 \le 3 \; \mathrm{d}$ and a 
  $\log$ scale for $t - t_0 > 3 \; \mathrm{d}$. The solid curve 
  represents the best-fit model from SALT2 
  (\S\ref{sec:classification}). The black ticks near the
  bottom of the panel show epochs of spectroscopic observations.
  \textit{Bottom}: Color curves of \abc. Each color curve is 
  represented with a different color, with each symbol corresponding 
  to a different instrument, as detailed in the legend. Magnitude 
  offsets applied to each color curve are also shown.
  }
  \label{fig:lightcurve}
\end{figure*}

During the spring of 2016, the iPTF survey observed the field of \abc\ every night in either the $g_\mathrm{PTF}$- or 
$R_\mathrm{PTF}$-band.\footnote{P48 observations of \abc\ is reported in the $g_\mathrm{PTF}$ and $R_\mathrm{PTF}$ filters throughout, which are similar to the SDSS $g'$ and $r'$ filters, respectively (see \citealt{2012PASP..124..854O} for details on PTF calibration). The correction from the $g_\mathrm{PTF}$ and $R_\mathrm{PTF}$ filters to SDSS $g'$ and $r'$ requires knowledge of the intrinsic source color (see Eqns.~1 and 2 in \citealt{2012PASP..124..854O}). The spectral diversity of SNe Ia in the days after explosion is poorly constrained, and as a result the color terms for \abc\ at these epochs are unknown. We proceed by assuming $g_\mathrm{PTF}$ and $R_\mathrm{PTF}$ calibration is on the AB system, which strictly speaking is incorrect, but this does not fundamentally alter any of our conclusions.} 
Survey observations were conducted with the
CFH12K camera \citep{2000SPIE.3965...58S} on the Palomar Observatory 
48-inch telescope (P48). Images were processed by the IPAC image
subtraction and discovery pipeline which subtracts off the background
galaxy light with stacked pre-SN images and performs forced
point-spread-function (PSF) photometry at the location of the SN 
\citep{2017PASP..129a4002M}. The
photometry is then calibrated to the PTF photometric catalog
\citep{2012PASP..124..854O}.

After discovery, photometric observations in the $g'$, $r'$ and $i'$
filters were obtained with the SED Machine 
(SEDm; Blagorodnova et al.\ 2017, in prep.) mounted on the Palomar Observatory 
60-inch telescope (P60). We utilized the Fremling Automated Pipeline \citep{2016A&A...593A..68F} to subtract galaxy light from the SEDm images using achival Sloan Digital Sky Survey (SDSS) images as a reference. This pipeline then performed forced-PSF photometry at the location of \abc, which is calibrated to the SDSS catalog \citep{2014ApJS..211...17A}.

Photometric observations in the \textit{BVg'r'i'} filters were conducted by the Las Cambres Observatory (LCO) 1-m
telescope network.  PSF photometry was measured on these images using
the \texttt{lcogtsnpipe} pipeline \citep{2016MNRAS.459.3939V}. The
\textit{BV} magnitudes are calibrated to the Fourth USNO CCD
Astrograph Catalog \citep{2013AJ....145...44Z}, and the \textit{g'r'i'}
magnitudes are calibrated to SDSS Data Release 6
\citep{2008ApJS..175..297A}.

The \textit{Swift} satellite observed \abc\ on 14 epochs, beginning 
$\sim$15 d pre-maximum light through $\sim$22 d post maximum. The SN 
flux is measured via aperture photometry on Ultraviolet-Optical
Telescope (UVOT) images via the usual procedures in 
\texttt{HEASoft}, including corrections for the coincident loss and 
aperture loss. The image counts
are converted to physical fluxes using the latest calibration
\citep{2011AIPC.1358..373B}. There are no pre-SN UVOT images at the 
SN location in the \textit{Swift} archive.  Visual inspection of the
UVOT images suggests negligible host-galaxy contamination in our 
UVOT flux measurements. No X-ray emission is detected from \abc\ by the \textit{Swift} X-ray Telescope (XRT).

The $g$-band discovery data and color curves of \abc\ are 
illustrated in Figure~\ref{fig:lightcurve}.  For convenience, 
magnitudes in 
\textit{all} filters are in the AB system with a zero point of 
$3631$\,Jy. As previously noted the color terms necessary to convert the $g_\mathrm{PTF}$ and $R_\mathrm{PTF}$ to the AB system are unknown and assumed to be zero.

Spectroscopic observations of \abc\ were taken with a variety
of telescopes and instruments over multiple epochs spanning from a
couple of days after explosion to two months after $B$-band maximum. 
An observing log is listed in Table \ref{tab:spec_obs_log}. The spectra were reduced using standard routines in \texttt{IDL}/\texttt{Python}. The optical
spectral evolution of \abc\ is illustrated in Figure
\ref{fig:spec_seq}, which excludes high-resolution Very Large Telescope (VLT) spectra for clarity.

\begin{deluxetable*}{cccccc}
  \tablecaption{Spectroscopic observations of \abc\ \label{tab:spec_obs_log}}
  \tablehead{
    \colhead{Observation MJD} & \colhead{SN phase} & \colhead{Telescope} &
    \colhead{Instrument} & \colhead{Wavelength Coverage (\AA)}
  }
  \startdata
  $57483.26$ & $-16.4$ & DCT & DeVeny\tablenotemark{1} & $3301$--$7499$ \\
  $57483.88$ & $-15.8$ & Gemini-North & GMOS\tablenotemark{2} & $3800$--$9200$ \\
  $57484.51$ & $-15.1$ & Keck-II & DEIMOS\tablenotemark{3} & $5500$--$8099$ \\
  $57486.51$ & $-13.1$ & Keck-II & DEIMOS\tablenotemark{3} & $5500$--$8099$ \\
  $57488.38$ & $-11.3$ & Keck-I & LRIS\tablenotemark{4} & $3055$--$10411$ \\
  $57489.51$ & $-10.1$ & LCOGT-2m & FLOYDS\tablenotemark{5} & $3301$--$8999$ \\
  $57490.40$ & $ -9.3$ & LCOGT-2m & FLOYDS\tablenotemark{5} & $3301$--$9999$ \\
  $57491.55$ & $ -8.1$ & LCOGT-2m & FLOYDS\tablenotemark{5} & $3300$--$9998$ \\
  $57492.20$ & $ -7.5$ & VLT & X-shooter\tablenotemark{6} & $3300$--$24550$ \\
  $57494.00$ & $ -5.7$ & VLT & UVES\tablenotemark{7} & \\
  $57503.32$ & $ +3.7$ & LCOGT-2m & FLOYDS\tablenotemark{5} & $3300$--$9999$ \\
  $57506.00$ & $ +6.3$ & NOT & ALFOSC\tablenotemark{8} & $3602$--$8098$ \\
  $57508.27$ & $ +8.6$ & LCOGT-2m & FLOYDS\tablenotemark{5} & $3301$--$9999$ \\
  $57518.42$ & $+18.8$ & Keck-I & LRIS\tablenotemark{4} & $3071$--$10208$ \\
  $57520.03$ & $+20.4$ & VLT & X-shooter\tablenotemark{6} & $3300$--$24789$ \\
  $57529.40$ & $+29.8$ & LCOGT-2m & FLOYDS\tablenotemark{5} & $4000$--$8998$ \\
  $57542.41$ & $+42.8$ & LCOGT-2m & FLOYDS\tablenotemark{5} & $4000$--$8998$ \\
  $57550.40$ & $+50.8$ & LCOGT-2m & FLOYDS\tablenotemark{5} & $4001$--$8999$ \\
  $57562.38$ & $+62.7$ & LCOGT-2m & FLOYDS\tablenotemark{5} & $4800$--$9300$ \\
  \enddata
  \tablenotetext{1}{The Deveny Spectrograph \citep{2014SPIE.9147E..2NB}}
  \tablenotetext{2}{The Gemini Multi-Object Sectrograph \citep{2004PASP..116..425H}}
  \tablenotetext{3}{DEep Imaging Multi-Object Spectrograph \citep{2003SPIE.4841.1657F}}
  \tablenotetext{4}{Low-Resolution Imaging Spectrometer \citep{1995PASP..107..375O}}
  \tablenotetext{5}{FLOYDS \url{https://lco.global/observatory/instruments/floyds}}
  \tablenotetext{6}{X-shooter \citep{XShooter}}
  \tablenotetext{7}{Ultraviolet and Visual Echelle Spectrograph \citep{2000SPIE.4008..534D}}
  \tablenotetext{8}{The Andalucia Faint Object Spectrograph and Camera \url{http://www.not.iac.es/instruments/alfosc}}
\end{deluxetable*}

\begin{figure*}[!htb]
  \centering
  \includegraphics[width=1.0\textwidth]{spectra.pdf}
  \caption{Observed spectral sequence of \abc. For clarity, offsets 
  have been applied to each 
  spectrum which are normalized by their median flux between 
  6,000 and 7,000$\,\textrm{\AA}$. The phase of each spectrum 
  relative to $T_{B_\mathrm{max}}$ is shown. Telluric absorption
  bands are grayed out. The narrow \ion{Na}{1}\,D absorption is also
  highlighted in orange.}
  \label{fig:spec_seq}
\end{figure*}

@@

A detailed study of the reddening towards \abc\ is presented in a 
companion paper (Ferretti et al.\ 2017, submitted). Briefly, the 
foreground Galactic extinction toward \abc\
is $E(B-V) = 0.0279 \, \mathrm{mag}$ \citep{2011ApJ...737..103S}. The high-resolution spectra presented in Ferretti et al.\ (2017) show multiple absorption components for both the \ion{Ca}{2}\,H$+$K and \ion{Na}{1}\,D doublets. While the equivalent width (EW) of these lines is quite large, implying a large amount of extinction (e.g., \citealt{2012MNRAS.426.1465P}), Ferretti et al.\ compare the evolution of \abc\ to the well-observed normal type Ia SN\,2011fe and find evidence for only a small amount of extinction. The empircal relation between the EW of \ion{Na}{1}\,D and extinction is known to have a large scatter, and thus, we instead adopt $E(B-V) = 0.05 \, \mathrm{mag}$ as the local extinction for \abc\ (Ferretti et al.\ 2017). For the remainder of our analysis we therefore adopt a total, Galactic $+$ host galaxy, line-of-sight extinction of $E(B-V) = 0.0779 \, \mathrm{mag}$.
